UKB Models

### 2023-10-06 ukb_mods.Rmd
region eda histograms

## Basic Plot Function

```{r}
mean.tbv.sim <- function(df, s, ageRange){
  tbv_list <- list()
  for (i in ageRange){
      tbv.df <- df %>%
        dplyr::filter(sex == s) %>%
        arrange(age_yrs) %>%
        dplyr::filter(age_yrs >= (i-0.005) & age_yrs <= (i+0.005)) %>%
        summarize(m = mean(TBV, na.rm = TRUE))
      tbv.val <-tbv.df$m
      tbv_list <- append(tbv_list, tbv.val)
  }
  return(unlist(tbv_list))
}

sim.data.ukb <- function(df){
  minAge <- min(df$age_days)
  maxAge <- max(df$age_days)
  ageRange <- seq(minAge, maxAge, 25)  # generate an age range with increments of 25 (since scaling in days)

  #sim data
  dataToPredictM <- data.frame(age_yrs=ageRange,
                               sex=c(rep(as.factor("Male"), length(ageRange))),
                               fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
                               sex.age_yrs=c(rep(0, length(ageRange))))
  dataToPredictM$TBV <- mean.tbv.sim(df, "Male", ageRange)

  dataToPredictF <- data.frame(age_yrs=ageRange,
                               sex=c(rep(as.factor("Female"), length(ageRange))),
                               fs_version=c(rep(Mode(df$fs_version), length(ageRange))),
                               study=c(as.factor(rep(Mode(df$study), length(ageRange)))),
                               sex.age_yrs=ageRange)
  dataToPredictF$TBV <- mean.tbv.sim(df, "Female", ageRange)

  # List of centiles for the fan plot
  desiredCentiles <- c(0.1, 0.25, 0.5, 0.75, 0.9)

  # return
  sim <- list(ageRange, dataToPredictM, dataToPredictF, desiredCentiles)
  names(sim) <- c("ageRange", "dataToPredictM", "dataToPredictF", "desiredCentiles")
  return(sim)
}

plot_singlestudy_centiles <- function(gamlssModel, phenotype, df, color_var)
{
  #use functions in GAMLSS repo plotting_functions.R to sim and predict
  sim <- sim.data.ukb(df) #simulate data
  pred <- centile_predict(gamlssModel, sim$dataToPredictM, sim$dataToPredictF, sim$ageRange, sim$desiredCentiles) #predict centiles

  #extract vals
  ages <- sim$ageRange
  age_col <- df$age_yrs
  male_peak_age <- pred$peak_age_M
  female_peak_age <- pred$peak_age_F
  unit_text <- "(years)"

  #plot!
  #sometimes df[,get()] works, sometimes not found...????
  yvar <- df[[phenotype]]
  color_col <- df[[color_var]]

  sampleCentileFan <- ggplot() +
    geom_point(aes(x=age_col, y=yvar, color=color_col), alpha=0.3) +
    scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) + 
    geom_line(aes(x=ages, y=pred$fanCentiles_M[[1]]), alpha=0.1) +
    geom_line(aes(x=ages, y=pred$fanCentiles_M[[2]]), alpha=0.3) +
    geom_line(aes(x=ages, y=pred$fanCentiles_M[[3]]), color="#40004BFF") +
    geom_line(aes(x=ages, y=pred$fanCentiles_M[[4]]), alpha=0.3) +
    geom_line(aes(x=ages, y=pred$fanCentiles_M[[5]]), alpha=0.1) +
    geom_line(aes(x=ages, y=pred$fanCentiles_F[[1]]), alpha=0.1) +
    geom_line(aes(x=ages, y=pred$fanCentiles_F[[2]]), alpha=0.3) +
    geom_line(aes(x=ages, y=pred$fanCentiles_F[[3]]), color="#00441BFF") +
    geom_line(aes(x=ages, y=pred$fanCentiles_F[[4]]), alpha=0.3) +
    geom_line(aes(x=ages, y=pred$fanCentiles_F[[5]]), alpha=0.1) +
    geom_point(aes(x=male_peak_age, y=pred$peak_M), color="#40004BFF", size=3) +
    geom_point(aes(x=female_peak_age, y=pred$peak_F), color="#00441BFF", size=3) +
    labs(title=deparse(substitute(gamlssModel))) +
    theme(legend.title = element_blank())+
    xlab(paste("Age at Scan", unit_text)) +
    ylab(deparse(substitute(phenotype)))

  print(sampleCentileFan)
}
```

>>>>

Look at variances
```{r}
plot.hist.by.sex <- function(data, pheno_list) {

  # Create an empty list to store the ggplot objects
  plots <- list()

  # Loop through the list of variables
  for (pheno in pheno_list) {
    # Check if the variable exists in the data frame
    if (!pheno %in% names(data)) {
      warning(paste("Variable '", pheno, "' not found in the data frame. Skipping..."))
      next
    }

    # Create a ggplot histogram for the variable
    plot <- ggplot(data, aes(x = .data[[pheno]], fill= sex, color=sex)) +
      geom_histogram(alpha=0.5, position="identity") +
      labs(title = paste("Histogram of", pheno), x = pheno, y = "Frequency")

    # Add the plot to the list
    plots[[pheno]] <- plot
  }

  # Return the list of ggplot objects
  return(plots)
}
```

center
```{r}
ukb_centered <- ukb_df_totals %>%
  group_by(sex) %>%
  mutate_if(is.numeric, funs(scale(., center=TRUE, scale=FALSE))) %>%
  ungroup()
```

```{r}
# Generate histograms (mean-centered w/in sex)
global_plots <- plot.hist.by.sex(ukb_centered, vol_list_global)

# Print or display the histograms
for (pheno in vol_list_global) {
  print(global_plots[[pheno]])
}
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_vol_plots <- plot.hist.by.sex(ukb_centered, vol_list_regions)

# Print or display the histograms
for (pheno in vol_list_regions) {
  print(region_vol_plots[[pheno]])
}
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_sa_plots <- plot.hist.by.sex(ukb_centered, sa_list)

# Print or display the histograms
for (pheno in sa_list) {
  print(region_sa_plots[[pheno]])
}
```

```{r}
# Generate histograms (mean-centered w/in sex)
region_ct_plots <- plot.hist.by.sex(ukb_centered, ct_list)

# Print or display the histograms
for (pheno in ct_list) {
  print(region_ct_plots[[pheno]])
}
```
